paper1_figures

### 2024-01-03 scoping_review_short.Rmd
intro + method + results okay

abstract: |
  Private ownership and commodification of resources are central aspects of the current capitalistic social organization. This has detrimental effects both on social and environmental levels. Commons represent an alternative to this system, therefore they could be the needed change toward a more sustainable society and economy. However, for a commons to be successful, its members must cooperate, i.e. incur a personal sacrifice for the collective good. Despite the importance of cooperation, there is a lack of studies that investigated this topic in the scientific field of the commons. On the other hand, cooperation has been extensively studied in social dilemma experiments, where personal and collective interests are at odds. Since social dilemma experiments and commons share some structural characteristics, we could gain insights into cooperation in commons and how to promote it thanks to the social dilemmas literature.
  The goal of the current scoping review is to identify the factors that impact cooperation in social dilemma situations, in order to better understand what could shape cooperation in commons. A key-word search and a subsequent selection led to the inclusion of 89 papers, which were used to identify 9 factors that influence cooperation both at the individual and group level: gender, social status, group identification, values and personality traits, trust, incentives, communication, social norms and anonymity. The influences that these elements can have on each-other are highlighted, showing that all factors should be taken into account when investigating cooperation in commons. Finally, the theoretical and practical implications of this work for commons' organization, as well as the review's limitations, are pointed out.

@@

<!-- change of the initial paragraph and of the definition -->

In recent years, the interest in alternative ways of social organization increased drastically due to the social and ecological crises we are facing [@Euler2018]. Commons could be an integral part of the path to attain a more sustainable and fair social organization [@Schmelzer2022]. In fact, they are a way of organizing resources beyond the classic private and public model [@Ostrom1990] and they are not organized around market competition [@Euler2019]. Commons express "a broader conception of property, referring to social goods [...] which a community, not the state or any individual, collectively owns, manages, and controls" [@Federici2011, p. 41]. Giving their characteristics, commons represent an option to the capitalistic system [@Euler2019] and its detrimental effects on natural resources and human well-being.

Commons can assume various forms, because "they can arise whenever a community decides it wishes to manage a resource in a collective manner, with a special regard for equitable access, use and sustainability" [@Bollier2014, p.19]. The variety of definitions can be resumed through three key points. First, commons allow to manage shared resources [@Ostrom1990] and to find solution for share needs of the population [@Euler2018; @Euler2019]. Second, commons are managed by a self-organized community of users [@Ostrom1990], i.e. the commoners. This community is a "social network of mutual aid, solidarity, and practices of human exchange that are not reduced to the market form" [@DeAngelis2003, p.1]. Moreover, commoners' organization is usually local, egalitarian, integrative, and fair [@Gidwani2011]. Third, the community's social practices are called "commoning" and, thanks to this process, commoners create institutions to govern the commons [@Euler2018; @Schmelzer2022]. Finally, although commons are often thought just as a shared resource, the other two aspects are also essential in defining them. Therefore, it is important to consider the commons as the integration of the resource, the community (i.e. the commoners) and its social practices (i.e. commoning) [@Bollier2014]. These definitions highlight the importance of the community as a part of the commons, and that is why it is essential to have a closer look at the people who form it.

More specifically, in order for a commons[^1] to be successful, its members have to cooperate, i.e to incur a personal sacrifice for the greater good [@Kollock1998]. Although it is essential, cooperation can be difficult to achieve due to two characteristics of the commons' social dynamic. First, people can believe that their individual contribution will not have a big impact, and this feeling of personal inefficacy discourages cooperation [@VanVugt1998]. Secondly, in a commons, every member faces a conflict between individual, short-term interests and collective, long-term ones [@Hardin1968]. Since acting for the common good requires a personal sacrifice, the temptation to pursue the individual interest is constantly present. Giving the role that commons could play in addressing the environmental and social problems, we need to understand the factors that affect cooperation, and how we can improve it. In fact, despite its importance, the reasons why individual choose to cooperate have not been extensively studied in commons scientific literature. At the same time, social dilemma experiments (i.e. artificial decision making situations, where personal and collective interests are at odds) allowed to identify several factors that could impact cooperation [e.g. @VanLange2013]. Since commons and social dilemmas share some structural characteristics [@Hardin1968; @Dawes1980], the aim of the current scoping review is to gain insight into cooperation in commons through social dilemma literature.

@@

@@

<!-- the whole structure has been changed according to Lilla's and Steffi's suggestions + Ostrom's principles explained more in details -->

One of the first paper about cooperation in commons was Hardin's "The tragedy of the commons" [-@Hardin1968]. In this article, Hardin argues that commons' members are rational beings who seek to maximize their personal interest by over-exploiting the common resource. This would inevitably lead to the destruction of the resource, unless an external institution regulates the access, or the resource is privatized. Therefore, according to Hardin, people benefiting from a common resource cannot successfully manage it without external interventions.

However, some years later, Elinor Ostrom's seminal work [-@Ostrom1990] falsified Hardin's predictions about people's inability to manage a common resource and their greed to maximize their self-interest. In fact, Ostrom observed several real-life commons, and showed that some of them were very successfully managed by the community of users without the intervention of an external authority. Based on her observations, she established seven design principles that are present in all the observed successfull commons, plus an eight one for more complex cases.[^2] By "design principle", Ostrom means "an essential element or condition that helps to account for the success of these institutions" [@Ostrom1990, p. 90].

[^2]: Since the eight principle is not relevant for the current paper, it will not be detailed.

The first principle posits that clearly defined boundaries are needed: it must be clear which is the shared resource, and who can benefit from it, otherwise, "no one knows what is being managed or for whom" (p. 91). Second, the rules must be congruent with the local conditions, i.e. they must reflect the specific attributs of the particular resource. Thirdly, individuals affected by the operational rules must be able to participate in modifying them. This ensure that the rules fit well to the local conditions and allows to promptly adapt them when needed. The fourth and fifth principles concern a monitoring and a graduate sanction system. Interestingly, these systems are successfully implemented by commoners them-selves, that can assume the roles of monitor or punisher, with no need of an external institution. In fact,

-   everybody has a self-interest in monitoring the others

-   no one will abuse of the sanctions because they know that they could be in the same situation so their are comprehensive + being caught has already a negative effect on reputation

1.  Monitoring + graduated sanctions (4 and 5) =\> these two systems are implemented by commoners them-selves, not by external institutions
    1.  Monitoring = low cost, because it is in everybody's self interest to monitor others + being a good monitor and catch somebody's infraction can make a person gain in prestige and status + material benefits if an infraction is caught (ex: part of the fine) + monitoring necessary to adopt a contingent strategy (if I cooperate, I want to be sure that the others do not exploit my work)
    2.  Not enourmous sanctions needed, because being caught is already negative for reputation
2.  Conflict resolution mechanism (6) =\> individuals following rules for a long period of time = inevitably some infractions, so it is essential to have somewhere to discuss when conflicts arise =\> give the possibility to explain the mistake, so that compliace to rules does not drop
3.  Minimal recognition of the right to organize (7) =\> external authorities (such governement) should not challenge the commoners' right to organize (if they do it, the rules of appropriators are not seen as valid and therefore more difficultly compliable)

These principles were not present in Hardin's example, where the hypothetical commoners made decisions in a context without local organization or established norms. Therefore, Hardin's scenario should not be considered a commons, but an open access regime [@Ostrom2008; @Bollier2014], and this can contribute to explain his negative predictions about cooperation.

At the same time, psychology is so far absent from the discussion

-   psychology is absent so far in the discussion, even though there are theories and experimental paradigms that are highliy relevant to the topic. On the theory side, there is....[the first paragraph with the different Theories] On the experimental side, there are classical social psychological experiments that...

Secondly, Hardin assumed, in line with the Rational Choice Theory, that all the commoners wanted to maximize their personal interest [@Wittel2023]. However, there are several theories less pessimistic than the RCT about human nature. For instance, the Greed Efficiency Fairness hypothesis posits that the individuals' greed to pursue their narrow interests is constrained by their desire to use the resource efficiently and to fairly allocate it between members [@Wilke1991]. Similarly, according to the Humanistic Rational Choice Theory, the desire to maintain basic social-psychological needs motivates people to cooperate [@DeCaro2021]. Moreover, cooperating can also be a self-presentation strategy [@VanVugt2010a], since showing altruistic behaviours has a positive effect on reputation [@Roberts1998].

Finally, the laboratory study of social dilemmas also largely contributed to the investigation of this topic, by allowing researchers to identify several factors that shape cooperation [e.g. @VanLange2013]. In general, these studies falsified RCT by demonstrating that, in social dilemmas, people cooperated more than predicted, and that their behaviour was impacted by variables that should not play a role according to the theory [@Ostrom1998]. Moreover, research based on different fields has convincingly demonstrated that people can engage in collective actions to sustainably manage shared resources [@Poteete2010a].

CONCLUSION: PSYCHOLOGY OFFER A SECOND EXPLANAITION OF WHY THERE IS NOT ALWAYS A TRAGEDY OF THE COMMONS = PEOPLE DO NOT ONLY WANT TO MAXIMISE THEIR PERSONAL INTEREST, AND SEVERAL "IRRATIONAL" FACTORS IMPACT THE WAY IN WHICH THEY INTERACT WITH OTHERS

@@

Commons and social dilemmas are deeply linked. First, they share the same basic conflict between personal and collective interests [@Hardin1968; @Dawes1980]. For this reason, commons can be considered as a particular case of social dilemma. In fact, @Kopelman2002 used the term of "commons dilemma" to define situations where a non-cooperative behaviour can seriously endanger the future of the shared resource. Moreover, commons have been classified by @Cumming2018 as "take-some dilemmas", a specific kind of social dilemma where people are tempted to increase their personal benefit by taking more than their fair share from a shared resource. Moreover, cooperation and defection in social dilemma laboratory experiments can predict people's behaviour in a real-life commons [e.g. @Reindl2019].

The current paper synthesizes existing research about cooperation in social dilemmas, to provide a set of factors that impact cooperative behaviours in commons. This work will therefore lay the foundations to implement experimental and quantitative studies in real-life commons, which, in turn, will allow to gain insights about cooperation beyond artificial laboratory settings. Other studies already used a social dilemma framework to study real-life commons [e.g. @Hartl2022]. While these studies usually focused on just one particular case, the current review is seeking to extend this approach by using social dilemma literature to gain a global understanding of cooperation in commons.

This scoping review has two main goals. The first is to identify factors on the individual level that can shape cooperation in commons. Moreover, since it has been widely demonstrated that commoners can successfully self-organize to govern a commons [@Ostrom1990; @Larsson2016], the second goal is to understand which strategies groups can implement to encourage their members to cooperate.

@@

To conduct this scoping review, we analyzed different scientific publications in response to the research questions. To gather the papers, we performed a Boolean search in Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", and "commons", and each of them was paired with each one of the following: "cooperation", "conflicts", "reciprocity", and "trust". We carried out the research during the month of August, 2023. According to @Munn2022, the type of source was not a selection criterion, therefore journal articles, conference papers and book chapters could be included, to have an overview as complete as possible about the existing literature on the topic. Moreover, we did not use any chronological or geographical limitations, once again to provide a broader vision. However, we implemented some selection criteria concerning the papers' methodology. First, we only selected papers written in English. Moreover, we focused on studies with human participants (or theoretical articles), so we excluded researches using an agent-based model. Secondly, among experimental or correlative studies, we only selected cases with adult participants.

Initially, 5210 papers emerged from the key phrases search. Deleting the duplicates reduced them to 4586. Based on titles and abstracts, we eliminated 4333 articles, leaving 253 texts to examine. After a full-text evaluation of the remaining papers, we selected 89 for the present review. Figure 1 resumes this selection process.

![Figure 1. Overview of the selection process.](images/figure1-01.png)

@@

The 89 included papers were published between 1984 and 2023, with a median of 2015. Almost all the selected texts were journal articles, with the only exception being a book chapter. Most of them were experimental studies (*N*=60), followed by literature review and meta-analysis (*N*=14) and correlative studies[^3] (*N*=14). The remaining paper was a qualitative study.

[^3]: Studies where at least one variable is manipulated by researchers are part of the category "experimental studies", while "correlative studies" refers to papers using quantitative analysis, but without manipulations.

@@

We identified two dimensions in the reviewed papers: setting (lab-based vs real-life) and context (social dilemma vs commons). There were papers in all four combinations of these dimentions, with the most represented category being "lab-based social dilemma" (*N*=75), namely highly controlled laboratory experiments where participants had to take cooperative decision. This was followed by "real-life social dilemma" (*N*=7), which investigated cooperation in different real-life situations with a social dilemma structure such as recycling behaviours [@Rompf2017] or use of public transportation [@VanLange1998]. Finally, "real-life commons" refers to studies of cooperation where participants were members of a commons in their everyday life, while "lab-based commons" refers to studies with a laboratory social dilemma paradigm, but with clear references to a concrete commons [e.g. @Mosler1993].

@@

Finally, most of the studies were conducted in Europe (*N*=40) (especially in Germany (*N*=12) and in the Netherlands (*N*=8)), followed by North America (*N*=20, of which *N*=19 in the USA) and Asia (*N*=7, of which *N*=6 in China). Moreover, six studies were set in more than one country, and fourteen were meta-analysis or literature reviews. The two remaining studies took place in Australia and Colombia, respectively.

A resume of the papers' description is provided in Table 1.

# Individual factors

Our analysis identified five individuals factors that impact cooperation: gender, social status, group identification, values and personality traits, and trust. In the following sections, we describe their principal effects on cooperation, and provide an overview of the moderators of each effect. More details about moderators are included in Appendix A.

## Gender

>>>>

abstract: |
  Private ownership and commodification of resources are central aspects of the current capitalistic social organization. Commons represent an alternative to this system, therefore they could be part of the solution to attain more sustainable society and economy. However, for a commons to be successful, its members must cooperate, i.e. incur a personal sacrifice for the collective good. Despite the importance of cooperation, there is a lack of studies that investigated this topic in the scientific field of the commons. On the other hand, cooperation has been extensively studied in social dilemma experiments, where personal and collective interests are at odds. Since social dilemma and commons share some structural characteristics, we could gain insights into cooperation in commons and how to promote it thanks to the social dilemmas literature.
  The goal of the current scoping review is to identify the factors that impact cooperation in social dilemma situations, in order to better understand what could shape cooperation in commons. A key-word search and a subsequent selection led to the inclusion of 89 papers, which were used to identify 9 factors that influence cooperation both at the individual and group level: gender, social status, group identification, values and personality traits, trust, incentives, communication, social norms and anonymity. The influences that these elements can have on each-other are highlighted, showing that all factors should be taken into account when investigating cooperation in commons. Finally, the theoretical and practical implications of this work for commons' organization, as well as the review's limitations, are pointed out.

@@

<!--# change of the initial paragraph and of the definition + adding the last paragraph with an overview of the goal of the paper -->

Commons allow to organize resources beyond the classic private and public model [@Ostrom1990] and they are not organized around market competition [@Euler2019]. They express "a broader conception of property, referring to social goods [...] which a community, not the state or any individual, collectively owns, manages, and controls" [@Federici2011, p. 41]. Giving their characteristics, commons represent an option to the capitalistic system, which is the main responsible for the current social and ecological crises [@Euler2019; @Wittel2023]. Therefore, they could be an integral part of the path to attain a more sustainable and fair social organization [@Schmelzer2022], and, for this reason, the interest in commons has recently increased [@Federici2019].

Commons can assume various forms, because "they can arise whenever a community decides it wishes to manage a resource in a collective manner, with a special regard for equitable access, use and sustainability" [@Bollier2014, p.19]. The variety of definitions can be resumed through three key points. First, commons allow to manage shared resources [@Ostrom1990] and to find solution for shared needs of the population [@Euler2018; @Euler2019]. Second, commons are managed by a self-organized community of users [@Ostrom1990], i.e. the commoners. This community is a "social network of mutual aid, solidarity, and practices of human exchange that are not reduced to the market form" [@DeAngelis2003, p.1]. Moreover, commoners' organization is usually local, egalitarian, integrative, and fair [@Gidwani2011]. Third, the community's social practices are called "commoning" and, thanks to this process, commoners create institutions to govern the commons [@Euler2018; @Schmelzer2022]. Finally, although commons are often thought just as a shared resource, the other two aspects are also essential in defining them. Therefore, it is important to consider the commons as the integration of the resource, the community (i.e. the commoners) and its social practices (i.e. commoning) [@Bollier2014]. For this reason, it is essential to have a closer look at the individuals who are part of commons.

More specifically, in order for a commons[^1] to be successful, its members have to cooperate, i.e to incur a personal sacrifice for the greater good [@Kollock1998]. Although it is essential, cooperation can be difficult to achieve due to two characteristics of the commons' social dynamic. First, people can believe that their individual contribution will not have a big impact, and this feeling of personal inefficacy discourages cooperation [@VanVugt1998]. Secondly, in a commons, every member faces a conflict between individual, short-term interests and collective, long-term ones [@Hardin1968]. Since acting for the common good requires a personal sacrifice, the temptation to pursue the individual interest is constantly present. Giving the role that commons could play in addressing the environmental and social problems, we need to understand the factors that affect cooperation, and how we can improve it.

@@

Despite its importance, cooperation has not been much investigated in commons scientific literature. At the same time, other disciplines such as psychology extensively studied it. More specifically, social dilemma experiments allowed to identify several factors that impact cooperation [e.g. @VanLange2013]. For this reason, the current scoping review is aimed at gaining insight into cooperation in commons through social dilemma literature. In the next section, we give an overview about the existing research about cooperation in commons, afterwards we will detail the research questions.

@@

<!--# the whole structure has been changed according to Lilla's and Steffi's suggestions + Ostrom's principles explained more in details -->

One of the first papers about cooperation in commons was Hardin's "The tragedy of the commons" [-@Hardin1968]. In this article, Hardin argues that commons' members are rational beings who seek to maximize their personal interest by over-exploiting the common resource. This would inevitably lead to the destruction of the resource, unless an external institution regulates the access, or the resource is privatized. Therefore, according to Hardin, people benefiting from a common resource cannot successfully manage it without external interventions.

However, research in different disciplines has convincingly demonstrated that people can engage in collective actions to sustainably manage shared resources, thus falsifying Hardin's predictions [@Poteete2010a]. Among these studies, the most important is Elinor Ostrom's seminal work [-@Ostrom1990], where she studied several real-life commons, some of them successfully managed by the community of users without the intervention of an external authority. Based on her observations, she established seven design principles that are present in all the successful commons, plus an eighth one for more complex cases.[^2]

[^2]: Since the eighth principle is not relevant for the current paper, it will not be detailed.

The first principle posits that boundaries are essential: it must be clear which is the shared resource, and who can benefit from it, otherwise, "no one knows what is being managed or for whom" (p. 91). Second, the rules must be congruent with the local conditions, i.e. they must reflect the specific attributes of the particular resource. Thirdly, individuals affected by the rules must be able to participate in modifying them. This ensure that the rules fit well to the local conditions and allows to promptly adapt them when needed. The fourth and fifth principles concern a monitoring and a graduate sanction system. Interestingly, these systems are successfully implemented by commoners themselves, with no need of an external institution. In fact, in order to know that the best strategy is to follow the rules, one must be sure that the others are also complying to them. For this reason, it is in everybody's self-interest to monitor other commoners. Similarly, the graduate sanction system can be easily self-regulated: even if punishing is usually profitable for the punisher, fines are naturally small for the first infraction and get higher only if the person persists in breaking the rules. In fact, since infractions are often due to extreme situations, punishers tend to be sympathetic because it could happen to them too. Linked to the previous principle, the sixth posits that people must have access to low-cost conflict resolution mechanisms, since, in a situation like a commons, disagreements between members are inevitable. Conflict resolution arenas give the rule-breaker the possibility to explain the mistake, and ensure that compliance to rules does not drop after every infraction. Finally, according to the seventh principle, external authorities should not challenge the commoners' right to self-organize and should recognize the commons' institutions.

These principles were not present in Hardin's example, where the hypothetical commoners made decisions in a context without local organization or established norms. Therefore, Hardin's scenario should not be considered a commons, but an open access regime [@Ostrom2008; @Bollier2014], and this can partly explain his negative predictions about cooperation.

At the same time, an individual perspective is so far absent from the discussion, although it provides further theoretical and experimental evidences against Hardin's predictions. First, there are several theories that offer an optimistic view of human ability to manage a shared resource. For instance, the Greed Efficiency Fairness hypothesis posits that the individuals' greed to pursue their narrow interests is constrained by their desire to use the resource efficiently and to fairly allocate it between members [@Wilke1991]. Similarly, according to the Humanistic Rational Choice Theory, cooperation can be seen as the rational option when the institution is accepted and perceived as legitimate, and when there is a strong group cohesion [@DeCaro2021]. Moreover, cooperating can also be a self-presentation strategy [@VanVugt2010a], since showing altruistic behaviours has a positive effect on reputation [@Roberts1998].

Secondly, on the experimental level, the laboratory study of social dilemmas also largely contributed to the investigation of cooperation, by identifying several factors that impact it [e.g. @VanLange2013]. A typical constellation in these experiments (often also referred as "public good games" or "common resource experiments") is that people face a conflict between personal and collective interest. Each individual obtains a higher personal payoff with a non-cooperative, self-interested behaviour than with a social, cooperative one, independently from what others do. At the same time, everybody's payoff is higher if all cooperate than if all defect [@Dawes1980]. Social dilemma experiments are very useful for gaining insight into cooperation, because they are easily implemented in a laboratory, and provide a standard experimental paradigm to test under which conditions people cooperate [@Balliet2011]. In general, these studies demonstrated that people are able to cooperate, and that their behaviour is impacted by variables that should not play a role from a rational point of view [@Ostrom1998].

Finally, to the best of our knowledge, there are no studies that investigated with experimental designs the causal effects of individual and group factors on people's willingness to cooperate in real-life commons. The aim of the current review is to lay the foundation to fill this gap in the literature.

@@

<!--# here I do not agree with Vincent and Lilla, who proposed to divide the three reasons to combine SD and commons based on the authors:1) the arguments are not completely separated: e.g. the fact that SD and commons share the basic conflict between personal and collective interest is the reason why commons are considered a specific kind of SD2) for the first argument I cited Hardin and Dawes not because they both said that SD and commons have the same characteristics, but because they defined the conflict in commons and SD, respectively (reading the two definitions is easy to make the link): for this reason, I would not consider them as the same "line" of research -->

Commons and social dilemmas are deeply linked. First, they share the same basic conflict between personal and collective interests [@Hardin1968; @Dawes1980]. For this reason, commons can be considered as a particular case of social dilemma. In fact, @Kopelman2002 used the term of "commons dilemma" to define situations where a non-cooperative behaviour can seriously endanger the future of a shared resource. Commons have also been classified by @Cumming2018 as "take-some dilemmas", a specific kind of social dilemma where people are tempted to increase their personal benefit by taking more than their fair share from a shared resource. Moreover, cooperation and defection in social dilemma laboratory experiments can predict people's behaviour in a real-life commons [e.g. @Reindl2019].

Based on this evidence, the current paper synthesizes existing research about cooperation in social dilemmas, to provide a set of factors that impact cooperative behaviours in commons. This work will therefore lay the foundations to implement experimental and quantitative studies in real-life commons, which, in turn, will allow to gain insights about cooperation beyond artificial laboratory settings.

Some works already went in a similar direction. For instance, @Hartl2022 used a social dilemma framework to study real-life commons. However, these authors focused on just one particular commons, while the current review provides more general results. Moreover, through a literature review, @Kopelman2002 identified several individual and situational factors that affect cooperation in commons dilemma. However, they restrained their review to papers published in peer-reviewed psychological journals. In the current review, we do not exclude any discipline beforehand, and this allows to provide a more complete overview of the factors impacting cooperation.

<!--# About the discussion with Steffi: Kopelman just focused on psychology, so it was interesting to insert it here to say that we are extending these results -->

In conclusion, the current review is seeking to extend what has already been done by using social dilemma literature to gain a global understanding of cooperation in commons. The first goal is to identify factors on the individual level that can shape cooperation in commons. The second goal is to understand which strategies groups can implement to encourage their members to cooperate, since it has been widely demonstrated that commoners can successfully self-organize to govern a commons [@Ostrom1990]. This leads to the three research questions:

1.  Which individual factors affect people's cooperation in social dilemmas?
2.  Which group factors affect people's cooperation in social dilemmas?
3.  Which moderating effects exist between the different factors?

<!--# Discussion with Steffi: goals = broader, questions = more specific for this reason in the goals I speak about commons (and strategies), while on the questions is SD (and factors) -->

@@

Due to the highly exploratory nature of the paper, a scoping review seemed more adapted to our goals than a systematic review. In fact, scoping reviews allow to collect the existing evidence on a specific topic, and to identify key factors linked to the main concept [@Munn2022].

To conduct this scoping review, we analyzed different scientific publications in response to the research questions. To gather the papers, we performed a Boolean search in Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", and "commons", and each of them was paired with each one of the following: "cooperation", "conflicts", "reciprocity", and "trust". To define these phrases, we screened key literature to identify (1) synonyms for "social dilemmas" and "commons", and (2) words that were frequently associated with "cooperation". We carried out the research during the month of August, 2023. Following the recommendations of @Munn2022, the type of source was not a selection criterion, therefore journal articles, conference papers and book chapters could be included, to have an overview as complete as possible about the existing literature on the topic. Moreover, we did not use any chronological or geographical limitations, once again to provide a broader vision.

Initially, 5210 papers emerged from the key phrases search. Deleting the duplicates reduced them to 4586. Based on titles and abstracts, we eliminated 4333 articles which were clearly not related to the topic[^3], leaving 253 texts to examine. During the full-text evaluation, we applied different criteria. First, we only selected papers written in English. Secondly, to be included, a paper must investigate one or more factors that impact cooperation in social dilemmas or in commons. Moreover, we focused on studies with human participants (or theoretical articles), so we excluded researches using an agent-based model. Finally, among experimental or correlative studies, we only selected cases with adult participants. Through the application of these criteria, we selected 89 papers for the present review. Figure 1 resumes this selection process.

[^3]: For instance, during the initial research, we found the article "International cooperation on financial market regulation" [@Abendschein2021] because, among its keywords, there were "public good" and "cooperation". However, the title already informed us that it would not allow us to answer the research questions.

<!--# add somewhere that qualitative studies are ok? -->

![Figure 1. Overview of the selection process.](images/figure1-03.png)

<!--# for the figure: "Papers", but not all of them are papers, maybe should I use "Texts"? -->

# Results

In this section we will firstly describe the characteristics of the included papers. Next, we will present the identified factors that impact cooperation, that can be at the individual or at the group level.

@@

+---------------------------+--------------+---------------+-------------+-------------------+---------------+--------+
|                           | Experimental | Correlational | Qualitative | Literature review | Meta-analysis | Total  |
+---------------------------+--------------+---------------+-------------+-------------------+---------------+--------+
| Lab-based social dilemmas | 52           | 7             | 1           | 7                 | 6             | 73     |
+---------------------------+--------------+---------------+-------------+-------------------+---------------+--------+
| Lab-based commons         | 5            | 0             | 0           | 0                 | 0             | 5      |
+---------------------------+--------------+---------------+-------------+-------------------+---------------+--------+
| Real-life social dilemmas | 2            | 5             | 0           | 0                 | 0             | 7      |
+---------------------------+--------------+---------------+-------------+-------------------+---------------+--------+
| Real-life commons         | 0            | 3             | 0           | 1                 | 0             | 4      |
+---------------------------+--------------+---------------+-------------+-------------------+---------------+--------+
| Total                     | 59           | 15            | 1           | 8                 | 6             | 89     |
+---------------------------+--------------+---------------+-------------+-------------------+---------------+--------+

: Table 1. Description of the selected papers.

Table 1 provides an overview of the 89 selected papers. The publication year ranges from 1984 to 2023, with a median of 2015. Almost all the selected texts were journal articles, with the only exception being a book chapter. Most of them were experimental studies (*N*=59), followed by correlative studies[^4] (*N*=14), and literature review and meta-analysis (*N*=14). The remaining paper was a qualitative study.

[^4]: Studies where at least one variable is manipulated by researchers are part of the category "experimental studies", while "correlative studies" refers to papers using quantitative analysis, but without manipulations.

@@

We identified two dimensions in the reviewed papers: setting (lab-based vs real-life) and context (social dilemma vs commons). There were papers in all four combinations of these dimensions, with the most represented category being "lab-based social dilemma" (*N*=73). In these experiments, participants had to take cooperative decision in a highly controlled laboratory setting. This was followed by "real-life social dilemma" (*N*=7), which investigated cooperation in different real-life situations with a social dilemma structure such as recycling behaviours [@Rompf2017] or use of public transportation [@VanLange1998]. Finally, "lab-based commons" (*N*=5) are studies with a laboratory social dilemma paradigm, but with clear references to a concrete commons [e.g. @Mosler1993], while "real-life commons" (*N*=4) refers to studies of cooperation where participants were members of a commons in their everyday life.

@@

Most of the studies were conducted in Europe (*N*=40) (especially in Germany (*N*=12) and in the Netherlands (*N*=8)), followed by North America (*N*=20, of which *N*=19 in the USA) and Asia (*N*=7, of which *N*=6 in China). Moreover, six studies were set in more than one country, and fourteen were meta-analysis or literature reviews. The two remaining studies took place in Australia and Colombia, respectively.

### Operationalization of cooperation

The way cooperation was operationalized was deeply connected to the study's setting. In fact, experiments based in laboratory often observed participants' distribution of resources between themselves and a group account to quantify their cooperation. The most used method (*N*=33) was to endow participants with an amount of a resource (often real or fictive money), and see which part they invested in a public account, from which the whole group could benefit. Otherwise, researchers observed which amount of the resource participants took for themselves from a collective pool (*N*=7), or how much they gave to an interaction partner (*N*=4).[^5] Another method often used in laboratory studies was to give participants the choice between cooperation and defection (*N*=18). Interestingly, in laboratory settings, words like "cooperation" or "defection" were rarely used during the instructions, to avoid responses dictated by social desirability.

[^5]: The use of real money allowed to provide a real consequence for the decisions taken in the laboratory. For the same reason, the studies that used fictive money often informed participants that their final sum would be converted in real money at the end of the experiment.

In studies with real-life settings, the operationalization of cooperation was more often based on observed behaviours (*N*=10). For instance, @Rompf2017 used the amount of recycling behaviours as proxy of cooperation, while in @VanLange1998 cooperation was the habits of taking public transports to commute.

## Individual factors

Our analysis identified five individuals factors that impact cooperation: gender, social status, group identification, values and personality traits, and trust. In the following sections, we describe their principal effects on cooperation, and provide an overview of the moderators of each effect presented in the reviewed papers. A resume of the different individual factors is provided in Table 2, while more details about moderators are included in Appendix A.

<!--# in the text: tables with the name of factor, the papers, the main effect and a list of moderatorsin Annexe: same table, but with more details for each moderator effect is it useful to put in both tables the analyzed papers for each factor?-->

+---------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+------------------------------------------------------------------+
| Individual factors                                            | Analyzed papers                                                                                                                                                                          | General effect on cooperation                                                                   | Moderators                                                       |
+===============================================================+==========================================================================================================================================================================================+=================================================================================================+==================================================================+
| Gender                                                        | *N*=11\                                                                                                                                                                                  | Ambiguous                                                                                       | -   Other participants' characteristics                          |
|                                                               | @Dorrough2019; @Peshkovskaya2017; @Balliet2011; @Spadaro2023; @Sell2009; @Gomez-Ruiz2020; @Colman2018; @Peshkovskaya2019; @Barrero-Amortegui2021; @Simpson2003; @Irwin2015               |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          |                                                                                                 | -   Group composition                                            |
|                                                               |                                                                                                                                                                                          |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          |                                                                                                 | -   Underlying motives                                           |
+---------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+------------------------------------------------------------------+
| Social status                                                 | *N*=6\                                                                                                                                                                                   | Ambiguous                                                                                       | -   Intra-class solidarity                                       |
|                                                               | @Aksoy2019; @Kingsley2016; @VanLange2013; @Chen2022; @Camera2020; @VanKlingeren2021                                                                                                      |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          |                                                                                                 | -   Group composition                                            |
+---------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+------------------------------------------------------------------+
| Group identification                                          | *N*=14\                                                                                                                                                                                  | Strongly identified individuals cooperate more with the in-group than low-identified ones       | -   Individual contributions                                     |
|                                                               | @Arora2016; @Noonan2016; @Gomez-Ruiz2020; @VanVugt2009; @Zhang2019; @Chen2007; @Wit1992; @Dorrough2015; @Aksoy2019; @Puurtinen2009; @Nockur2020; @Kramer1984; @Jackson2011; @Jackson2012 |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          |                                                                                                 | -   Intergroup competition                                       |
|                                                               |                                                                                                                                                                                          |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          |                                                                                                 | -   Super-ordinate identity                                      |
+---------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+------------------------------------------------------------------+
| Social Value Orientation and other values/personality factors | *N*=5 (SVO)\                                                                                                                                                                             | Pro-social individuals cooperate more than pro-self                                             | -   Incentives                                                   |
|                                                               | @Balliet2009; @Emonds2011; @VanLange1998; @DeKwaadsteniet2008; @Lu2019                                                                                                                   |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          | Pro-environmental values are positively correlated to cooperation                               | -   Uncertainty about the number of members                      |
|                                                               | *N*=9 (other factors)\                                                                                                                                                                   |                                                                                                 |                                                                  |
|                                                               | @Sussman2016; @Fosgaard2019; @Kieslich2014; @Hilbig2018; @Kocher2017; @Schlosser2018; @Liu2009a; @Probst1999; @Chen2007                                                                  | Left-wingers are more cooperative than right-wingers                                            | -   Deliberation                                                 |
|                                                               |                                                                                                                                                                                          |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          | Individuals scoring high in humility and honesty cooperate more than those scoring low          | -   Inter- vs independent self                                   |
|                                                               |                                                                                                                                                                                          |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          | Individuals with high self-control, or with a high level of justice sensitivity, cooperate more | -   Perceived social norms                                       |
+---------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+------------------------------------------------------------------+
| Trust                                                         | *N*=10\                                                                                                                                                                                  | People with high levels of trust cooperate more than people with low levels                     | -   Degree of conflict between personal and collective interests |
|                                                               | @Balliet2013; @Lubke2021; @VanLange2013; @Kocher2015; @Smith2013; @Chaudhuri2002; @Oyediran2018; @Sturm2019; @Franzen2019; @Rompf2017                                                    |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          |                                                                                                 | -   Uncertainty about others' intentions                         |
|                                                               |                                                                                                                                                                                          |                                                                                                 |                                                                  |
|                                                               |                                                                                                                                                                                          |                                                                                                 | -   Perceived costs and benefits of the cooperative behaviour    |
+---------------------------------------------------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+------------------------------------------------------------------+

: Table 2. Resume of individual factors.
